\MakeUppercase{Graph Product Structure for Non-Minor-Closed Classes}

### 2022-03-11 k-planar.tex
Pfffft
not man yupdates

\begin{proof}[Proof of \cref{kPlanarProduct}]
	Let $G$ be a $k$-plane graph.  We will assume, for ease of exposition, that any point $p\in\R^2$ is involved in at most one crossing $(p,vw,xy)$ of $G$. This assumption is justified since it can be enforced by a slight deformation of the edges of $G$ and the resulting (deformed) graph is also $k$-plane.

@@

In the following, we will assume that no two edges incident to a common vertex of the $1$-plane graph $G$ cross each other since, in a 1-plane graph, such a crossing can always be removed by a local modification to obtain an isomorphic 1-plane graph in which the two edges do not cross.\footnote{While this is true for 1-plane graphs it is not true for $k$-plane graphs with $k\ge 3$; the uncrossing operation can increase the number of crossings on a particular edge from $k$ to $2(k-1)$.}  The \defin{faces} of $G$ are the connected components of $\R^2\setminus \bigcup_{vw\in E(G)} vw$, where we treat each edge $vw$ of $G$ as the image of a simple curve $f_{vw}:[0,1]\to\R^2$.  We say that a vertex $v\in V(G)$ appears on a face $F$ if $v$ is contained in the closure of $F$.

Let $G$ be a $1$-plane graph.  
% Then $G$ has $O(V(G)$ edges since the edges of $G$ can be partitioned into two planar graphs by placing one edge from each crossing pair into a second set.
% \footnote{We may assume that $G$ has no parallel edges, since removing all but one copy from each set of parallel edges does not change the set of adjacent pairs of vertices in $G$, so any supergraph that contains}
We make $G$ into an \emph{edge-maximal} $1$-plane multigraph, by repeating the following operation:  If two vertices $v$ and $w$ of $G$ appear on a common face $F$ and there is no edge $vw\in E(G)$ contained in the boundary of $F$, then add $vw$, embedded in the face $F$.  This may introduce parallel edges into $G$, but no two such edges will appear on the boundary of a common face.  
% This increases the number of edges of $G$, by $O(|V(G)|)$ since these newly added edges do not cross any other edges of $G$.

@@

To understand the structure of $G$, it is helpful to consider the plane graph $G_0$ obtained by replacing each pair of edges $vw$ and $xy$ involved in a crossing $(p,vw,xy)$ with four edges $vp$, $wp$, $xp$, and $yp$ meeting at a newly added \emph{dummy vertex} $p$. Let $F$ be a face of $G_0$ and let $W:=w_0,\ldots,w_{r-1}$ be the facial walk around $F$. In the following paragraphs, subscripts on the vertices in $W$ are implicitly taken modulo $r$. By edge maximality, if $W$ contains only non-dummy vertices, then it contains exactly three vertices and $F$ is bounded by three edges of $G$.

If $W$ contains a dummy vertex $w_i$ then neither $w_{i-1}$ nor $w_{i+1}$ is a dummy vertex since if $w_{i-1}$ (respectively $w_{i+1}$) were a dummy vertex then the edge of $G$ that contains $w_{i-1}w_i$ (respectively $w_iw_{i+1}$) would be involved in at least two crossings.  Therefore each dummy vertex $w_i$ of $W$ is surrounded by two non-dummy vertices $w_{i-1},w_{i+1}\in V(G)$.  Since $w_i$ has degree $4$ in $G_0$, $w_{i-1}\neq w_{i+1}$.  Since $G$ is edge-maximal the edge $w_{i-1}w_i$ is an edge of $G$ on the boundary of $F$.  Therefore, if $F$ has a dummy vertex $w_i$ in its facial walk $W$, then $W=w_{i-1},w_i,w_{i+1}$ and $F$ is bounded by three edges of $G_0$, each of which is contained in a different edge of $G$.

@@

Our proof of \cref{1-planar} follows quickly from the following technical lemma, which is an extension of the analogous result for plane graphs \cite{DJMMUW20}.
\begin{lem}
	\label{induction} The setup:
	\begin{compactenum}
		\item Let $G$ and $G'$ be defined as above.
		\item Let $T$ be a BFS spanning tree of $G'$ rooted at some vertex $r$.
		\item For every integer $j\ge 0$, let $L_j=\{v\in V(G):\dist_T(r,v)=j\}$.
		\item Let $F$ be a cycle in $G'$ with $r$ in the exterior of $F$ and such that
		\begin{compactenum}
			\item No edge of $F$ is crossed by any edge of $G$; and
			\item $V(F)$ can be partitioned into $P_1,\ldots,P_k$, for some $k\in\{1,2,3\}$ such that for each $i\in\{1,\ldots,k\}$,
			\begin{compactenum}
				\item $F[P_i]$ is a path; and
				\item $|V(P_i)\cap L_j| \le 3$ for every integer $j\ge 0$.
			\end{compactenum}
		\end{compactenum}
% \note{DW}{Should the vertices already grabbed by $P_1,P_2,P_3$ be mentioned in the setup? The new paths $Q_1,Q_2,Q_3$ need to avoid the vertices already grabbed by $P_1,P_2,P_3$. Basically, I am concerned about what happens when distinct vertical paths in $T$ want to grab the same vertex.} \note{PM}{No, at this point we're only promising an $H$-partition of $N$, the part of $G$ contained in the uncrossed cycle $F$.}
		\item Let $N$ and $N'$ be the subgraphs of $G$ and $G'$ consisting only of those edges and vertices contained in $F$ or the interior of $F$.
	\end{compactenum}
	Then $N$ has an $H$-partition $\PP=\{S_x\cup\{v_x\} : x\in V(H)\}$ such that:
	\begin{compactenum}
		\item $H$ is planar;
		\item for every integer $j\ge 0$ and every $x\in V(H)$, $|S_x\cap L_j|\le 3$;
		\item for each $i\in\{1,\ldots,k\}$, there exists some $x_i\in V(H)$ such that $P_i=S_{x_i}$; 
%		\note{DW}{Should this be $P_i\subseteq S_{x_i}$? (since the path grabs other vertices into its part)}\note{PM}{No, for the same reason as above.}
 and
		\item $H$ has a tree-decomposition in which every bag has size at most 4 and such that some bag contains $x_1,\ldots,x_k$.
	\end{compactenum}
\end{lem}

@@

  If $k\in\{1,2\}$ then we let $\tau:=v_1,\ldots,v_p$ be an inner face of $N'$ that contains at least one edge of $F$ on its boundary.  If $k=3$ then assign each vertex $v$ of $N'$ a colour $\alpha(v)\in\{1,\ldots,3\}$ as follows:  Let $w$ be the first vertex of $F$ encountered on the path in $T$ from $v$ to the root of $T$.  (This implies that $w=v$ when $v$ is a vertex of $F$.)  Since $w\in V(F)$, $w\in P_i$ for some $i\in\{1,\ldots,k\}$ and we define $\alpha(v):=i$.  Now add edges to the inner faces of $N'$ to obtain a supergraph $N''$ in which each inner face has three $3$ edges on its boundary.
  By Sperner's Lemma there exists an inner face $\tau'':=v_1,v_2,v_3$ of $N''$ such that $\alpha(v_i)=i$ for each $i\in\{1,2,3\}$.  Let $\tau:=v_1,\ldots,v_p$ be the face of $N'$ that contains $\tau''$.  (Note that $p\in\{3,4\}$.)

@@

\referee{2}{Page 16.  At the end of Page 16, I do not see why 'every pair of edges cross at most once.'  Without further assumptions, it seems as if two paths in the path system could intersect several times.  Therefore, by drawing 'each edge vw of G alongside $P_{vw}$ in $G_0$', there may be two edges which cross several times.  It seems this can be fixed by choosing the path system carefully by 'uncrossing' paths which intersect at more than two internal vertices.}

\note{DW}{I have revised the proof to address this error. I couldn't think of a way to do this without introducing balls and channels (and without vigorous hand-waving). Is it readable?}

@@

\citet{PS21} and \citet{DFMS21} use \cref{ps} and \cref{PlanarProduct}(b), respectively, to show that $\chi_p(G)$ is polynomial in $p$ when $G$ is planar or of bounded Euler genus.  We use the following lemma due to \citet[Lemma~15]{PS21}.

\begin{lem}[\citep{PS21}]
\label{p-centered-treewidth}
  Every graph $H$ of treewidth at most $t$ has $\chi_p(H)\leq \binom{p+t}{t}$.
\end{lem}

\note{DW}{In the journal version \citet{DFMS21} (under the insistence of the referee :-) ) proved that $\chi_p(H_1\boxtimes H_2) \leq \chi_p(H_1) \cdot \chi(H_2^p)$ where $H_2^p$ is the $p$-th power of $H_2$. Apply this with $H_2:= P \boxtimes K_\ell$. Then $H_2^p= P^p\boxtimes K_\ell$ and $\chi(H_2^p)=(p+1)\ell$. So
	$\chi_p(H\boxtimes P \boxtimes K_\ell) \leq \chi_p(H_1) \cdot (p+1)\ell$. So I suggest we replace the next paragraph by 	``\citet{DFMS21} proved that
$\chi_p(H \boxtimes J) \leq \chi_p(H) \cdot \chi(J^p)$ for all graphs $H$ and $J$, which implies the following lemma'' and delete the proof. Okay?}

The following lemma is implicitly in the work of \citet[Proof of Theorem~2.1]{DFMS21}. We include the proof for completeness.

\begin{lem}[\citep{DFMS21}]
\label{p-centered}
If $G\subseteq H\boxtimes P \boxtimes K_\ell$ then $\chi_p(G)\le \ell (p+1)\, \chi_p(H)$.
\end{lem}

\note{DW}{For consistency with \cref{qn}, I suggest we combine \cref{p-centered,p-centered-treewidth}, and just write, ``For every graph $H$ of treewidth at most $t$ and for every path $P$, if $G\subseteq H\boxtimes P \boxtimes K_\ell$ then
	\[\chi_p(G)\le \ell (p+1)\, \chi_p(H) \leq \ell (p+1) \tbinom{p+t}{t}.\]}

\begin{proof}
	By \cref{PartitionProduct}, $G$ has an $H$-partition $(\mathcal{L}=\langle L_0,L_1,\ldots\rangle, \PP=(B_x:x\in V(H))$ with layered width at most $\ell$. Use a product colouring $\phi:V(G)\to \{1,\ldots,\ell\}\times\{0,\ldots,p\}\times\{1,\ldots,\chi_p(H)\}$.  For each integer $i\ge 0$ and each $x\in V(H)$, assign the colour $\phi(v):=(\alpha(v),\beta(v),\gamma(v))$ to each vertex $v\in L_i\cap B_x$ such that:
   \begin{compactenum}
     \item $\alpha(v)$ is unique among $\{\alpha(w): w\in L_i\cap B_x\}$, which is possible
     since $|L_i\cap B_x|\le \ell$,
     \item $\beta(v)= i\bmod (p+1)$, and
     \item $\gamma(v)=\gamma(x)$ where $\gamma:V(H)\to\{1,\ldots,\chi_p(H)\}$ is a $p$-centered colouring of $H$.
   \end{compactenum}
 To show this is a $p$-centered colouring, consider some connected subgraph $X\subseteq G$.

 First suppose that there exists $v,w\in V(X)$ with $v\in L_i$ and $w\in L_j$ with $j-i\ge p$. Since $X$ is connected, $X$ contains a path from $v$ to $w$.  By the definition of layering, this path contains at least one vertex from $L_{i'}$ for each $i'\in\{i,i+1,\ldots,j\}$. Therefore, $|\{\beta(v'):v'\in V(X)\}|\ge j-i+1 > p$, so $X$ receives more than $p$ distinct colours.

 Otherwise, $V(X)\subseteq L_{i}\cup\cdots\cup L_{i+s}$ for some $s<p$.  Let $H':=H[\{x\in V(H):B_x\cap V(X)\neq\emptyset]$.  If $|\{\gamma(x):x\in V(H')\}| > p$ then $|\{\gamma(v):v\in V(X)\}|> p$ so $|\{\phi(v):v\in V(X)\}|> p$ and we are done.  Otherwise, since $\gamma$ is a $p$-centered colouring of $H$, there must exist some $x\in V(H')$ such that $\gamma(x)\neq\gamma(y)$ for every $y\in V(H')\setminus\{x\}$.
 For any $v,w\in B_x$ with $v\neq w$, either $v,w\in L_{i'}$ for some $i'\in\{i,i+1,\ldots,i+s\}$ in which case $\alpha(v)\neq\alpha(w)$; or $v\in L_{i'}$ and $w\in L_{i''}$ with $|i'-i''|< p$, in which case $\beta(v)\neq\beta(w)$. Therefore every vertex $v\in B_x$ receives a colour $\phi(v)$ distinct from every colour in $\{\phi(z):z\in X\setminus\{x\}\}$. Therefore, every vertex in $B_x$ receives a colour distinct from every other vertex in $X$.
 \end{proof}

\referee{2}{Page 18.  Replace '$|i'-i''| < p$' by  '$0< |i'-i''| < p$'.}

\cref{p-centered,p-centered-treewidth,kPlanarProduct,1-planar,gkPlanarProduct} immediately imply the following results, for every $p\ge 2$:

@@

Theorems~\ref{PlanarProduct}(b),  \ref{ShortcutProduct} and  \ref{GenusProduct}(b) and \cref{MapShortcut,qn,p-centered,p-centered-treewidth,non-repetitive} imply the following result.

@@

Theorems~\ref{PlanarProduct}(b), \ref{GenusProduct}(b) and \cref{ShortcutProduct} and  \cref{DrawG,qn,p-centered,p-centered-treewidth,non-repetitive,PowerShortcut} imply the following result, which with $g=0$ implies  \cref{kPowerBasic} in the introduction.

@@

\cref{qn,p-centered,p-centered-treewidth,non-repetitive,MinorFreeDegree,PowerProduct} imply the following analogous result for powers of graphs in any minor-closed class with bounded maximum degree.

@@

\cref{nearest-neighbour,kPlanarProduct,qn,p-centered,p-centered-treewidth} imply:

@@

\begin{lem}
	\label{NewLemma}
Let $G_0$ be a 2-connected plane multigraph. Let $T$ be a spanning tree of $G_0$ rooted at some vertex $r$ on the boundary of the outerface of $G_0$. Let $C = [P_1,\dots,P_k]$ be a cycle in $G_0$ with $r$ in the exterior of $C$, where $k\in\{1,2,3\}$ and $P_1, \dots, P_k$ are bipods in $G_0$. Let $G:=G_0\blah{C}$. Then $G$ has a partition $\PP$ into tripods with respect to $T$ such that:
\begin{compactitem}
\item $P_1,\dots,P_k\in\PP$,
\item $G / \PP$ has a tree-decomposition $\mathcal{T}$ of width at most 3 such that:
\begin{compactenum}[(i)]
	\item some bag of $\mathcal{T}$ contains $P_1,\dots,P_k$, and 
	\item for every facial cycle $F$ of $G$ there is a bag $B$ of $\mathcal{T}$ such that $V(F)\subseteq \bigcup_{X\in B} X$.
\end{compactenum}
\note{DW}{We need to replace (ii) by ``for every facial cycle $F$ of $G$ there is a bag $B$ of $\mathcal{T}$ such that every part of $\PP$ that intersects $F$ is in $B$''. Curiously, this implies (i) (since $C$ is a facial cycle of $G$).}
\end{compactitem}
\end{lem}

@@

First suppose that $C$ is a 2-cycle. Thus $k\leq 2$. Let $e$ be one of the edges in $C$. Let $F$ be the internal facial cycle of $G$ that includes $e$. Since each facial cycle of $G_0$ has length at least 3, $F$ consists of a non-empty path $P$ plus the edge $e$. Let $C'$ be obtained by replacing $e$ in $C$ by $P$. Let $G':=G\langle{C'}\rangle$. Thus $G'$ has one fewer face than $G$. Note that $C'=[P_1,\dots,P_k,P]$ (\note{DW}{or it can be written that way?}).  By induction, $G'$ has a partition $\PP$ into tripods with respect to $T$ such that:
\begin{compactitem}
	\item $P_1,\dots,P_k,P\in\PP$,
	\item $G / \PP$ has a tree-decomposition $\mathcal{T}$ of width at most 3 such that:
	\begin{compactenum}[(i)]
	\item some bag of $\mathcal{T}$ contains $P_1,\dots,P_k,P$, and 
	\item for every facial cycle $F'$ of $G\blah{C'}$ there is a bag $B$ of $\mathcal{T}$ such that $V(F)\subseteq \bigcup_{X\in B} X$.
	\end{compactenum}
\end{compactitem}
The bag of $\mathcal{T}$ that contains $P_1,\dots,P_k,P$ satisfies property (ii) for $F$. 
Thus, $\PP$ and $\TT$ satisfy the desired properties for $G$. Now assume that $|V(C)|\geq 3$.

@@

Consider $j\in\{1,\dots,m\}$. Let $G_j$ be the subgraph of $G$ consisting of $Z_j$ and all the edges and vertices of $G$ in the interior of $Z_j$. The outerface of $G_j$ is $Z_j$. Observe that for some $i\in\{1,2,3\}$, we have
$Z_j=[R_{i,j}^+,R_{i+1,j}^-,S_{i,j}]$, where 
$R_{i,j}^+$ is some subpath of $R_i^+$,
$R_{i+1,j}^-$ is some subpath of $R_{i+1}^-$, and
$L_{i,j}$ is the path $Z_j-V(C)$ (which is contained in $Y$). \note{DW}{Add a figure of a good case and a weird case.}
Since $R_i^+$, $R_{i+1}^-$ and $L_i$ are bipods, 
$R_{i,j}^+$, $R_{i+1,j}^-$ and $L_{i,j}$ are bipods. 
By construction, $v_{i-1}$ is not in $G_j$ (where $v_0=v_3$). 
Thus $V(G_j)\setminus V(Z_j) \subsetneq V(G)\setminus V(C)$, and we may apply induction to $G_j$.  Thus $G_j$ has a partition $\PP_j$ into tripods with respect to $T$ such that:
\begin{compactitem}
	\item $R_{i,j}^+, R_{i+1,j}^-, L_{i,j}\in\PP_j$, 
	\item the quotient $G_j / \PP_j$ has a tree-decomposition $\mathcal{T}_j=\{B_x:x\in V(D_j)\}$ of width at most 3 such that: 
	\begin{compactitem}
		\item some bag $B_{u_j}$ of $\mathcal{T}_j$ contains $R_{i,j}^+, R_{i+1,j}^-, L_{i,j}$, and 
		\item for every facial cycle $F'$ of $G_j$ there is a bag $B$ of $\mathcal{T}_j$ such that $V(F')\subseteq \bigcup_{X\in B} X$.
	\end{compactitem}
\end{compactitem}

>>>>

\begin{proof}[Proof of \cref{kPlanarProduct}]
	Let $G$ be a $k$-plane graph.  We assume, for ease of exposition, that any point $p\in\R^2$ is involved in at most one crossing $(p,vw,xy)$ of $G$. This assumption is justified since it can be enforced by a slight deformation of the edges of $G$ and the resulting (deformed) graph is also $k$-plane.

@@

In the following, we assume that no two edges incident to a common vertex of the $1$-plane graph $G$ cross each other since, in a 1-plane graph, such a crossing can always be removed by a local modification to obtain an isomorphic 1-plane graph in which the two edges do not cross.\footnote{While this is true for 1-plane graphs it is not true for $k$-plane graphs with $k\ge 3$; the uncrossing operation can increase the number of crossings on a particular edge from $k$ to $2(k-1)$.}  The \defin{faces} of $G$ are the connected components of $\R^2\setminus \bigcup_{vw\in E(G)} vw$, where we treat each edge $vw$ of $G$ as the image of a simple curve $f_{vw}:[0,1]\to\R^2$.  We say that a vertex $v\in V(G)$ appears on a face $F$ if $v$ is contained in the closure of $F$.

Let $G$ be a $1$-plane graph.  
% Then $G$ has $O(V(G)$ edges since the edges of $G$ can be partitioned into two planar graphs by placing one edge from each crossing pair into a second set.
% \footnote{We may assume that $G$ has no parallel edges, since removing all but one copy from each set of parallel edges does not change the set of adjacent pairs of vertices in $G$, so any supergraph that contains}
We make $G$ into an \defin{edge-maximal} $1$-plane multigraph, by repeating the following operation:  If two vertices $v$ and $w$ of $G$ appear on a common face $F$ and there is no edge $vw\in E(G)$ contained in the boundary of $F$, then add $vw$, embedded in the face $F$.  This may introduce parallel edges into $G$, but no two such edges will appear on the boundary of a common face.  
% This increases the number of edges of $G$, by $O(|V(G)|)$ since these newly added edges do not cross any other edges of $G$.

@@

To understand the structure of $G$, it is helpful to consider the plane graph $G_0$ obtained by replacing each pair of edges $vw$ and $xy$ involved in a crossing $(p,vw,xy)$ with four edges $vp$, $wp$, $xp$, and $yp$ meeting at a newly added \defin{dummy vertex} $p$. Let $F$ be a face of $G_0$ and let $W:=w_0,\ldots,w_{r-1}$ be the facial walk around $F$. In the following paragraphs, subscripts on the vertices in $W$ are implicitly taken modulo $r$. By edge maximality, if $W$ contains only non-dummy vertices, then it contains exactly three vertices and $F$ is bounded by three edges of $G$.

If $W$ contains a dummy vertex $w_i$, then neither $w_{i-1}$ nor $w_{i+1}$ is a dummy vertex since if $w_{i-1}$ (respectively $w_{i+1}$) were a dummy vertex then the edge of $G$ that contains $w_{i-1}w_i$ (respectively $w_iw_{i+1}$) would be involved in at least two crossings.  Therefore each dummy vertex $w_i$ of $W$ is surrounded by two non-dummy vertices $w_{i-1},w_{i+1}\in V(G)$.  Since $w_i$ has degree $4$ in $G_0$, $w_{i-1}\neq w_{i+1}$.  Since $G$ is edge-maximal the edge $w_{i-1}w_i$ is an edge of $G$ on the boundary of $F$.  Therefore, if $F$ has a dummy vertex $w_i$ in its facial walk $W$, then $W=w_{i-1},w_i,w_{i+1}$ and $F$ is bounded by three edges of $G_0$, each of which is contained in a different edge of $G$.

@@

Our proof of \cref{1-planar} follows quickly from the following technical lemma, which is an extension of the analogous result for plane graphs \cite{DJMMUW20}.
\begin{lem}
	\label{induction} The setup:
	\begin{compactenum}
		\item Let $G$ and $G'$ be defined as above.
		\item Let $T$ be a BFS spanning tree of $G'$ rooted at some vertex $r$.
		\item For every integer $j\ge 0$, let $L_j=\{v\in V(G):\dist_T(r,v)=j\}$.
		\item Let $F$ be a cycle in $G'$ with $r$ in the exterior of $F$ and such that
		\begin{compactenum}
			\item No edge of $F$ is crossed by any edge of $G$; and \note{DW}{This can be deleted since $F$ is in $G'$, right?}
			\item $V(F)$ can be partitioned into $P_1,\ldots,P_k$, for some $k\in\{1,2,3\}$ such that for each $i\in\{1,\ldots,k\}$,
			\begin{compactenum}
				\item $F[P_i]$ is a path; and
				\item $|V(P_i)\cap L_j| \le 3$ for every integer $j\ge 0$.
			\end{compactenum}
		\end{compactenum}
% \note{DW}{Should the vertices already grabbed by $P_1,P_2,P_3$ be mentioned in the setup? The new paths $Q_1,Q_2,Q_3$ need to avoid the vertices already grabbed by $P_1,P_2,P_3$. Basically, I am concerned about what happens when distinct vertical paths in $T$ want to grab the same vertex.} \note{PM}{No, at this point we're only promising an $H$-partition of $N$, the part of $G$ contained in the uncrossed cycle $F$.}
		\item Let $N$ and $N'$ be the subgraphs of $G$ and $G'$ consisting only of those edges and vertices contained in $F$ or the interior of $F$.
	\end{compactenum}
	Then $N$ has an $H$-partition $\PP=\{S_x\cup\{v_x\} : x\in V(H)\}$ such that:
	\begin{compactenum}
		\item $H$ is planar;
		\item for every integer $j\ge 0$ and every $x\in V(H)$, $|S_x\cap L_j|\le 3$;
		\item for each $i\in\{1,\ldots,k\}$, there exists some $x_i\in V(H)$ such that $P_i=S_{x_i}$; 
%		\note{DW}{Should this be $P_i\subseteq S_{x_i}$? (since the path grabs other vertices into its part)}\note{PM}{No, for the same reason as above.}
 and
		\item $H$ has a tree-decomposition in which every bag has size at most 4 and such that some bag contains $x_1,\ldots,x_k$.
	\end{compactenum}
\end{lem}

@@

  If $k\in\{1,2\}$ then we let $\tau:=v_1,\ldots,v_p$ be an inner face of $N'$ that contains at least one edge of $F$ on its boundary.  If $k=3$ then assign each vertex $v$ of $N'$ a colour $\alpha(v)\in\{1,\ldots,3\}$ as follows:  Let $w$ be the first vertex of $F$ encountered on the path in $T$ from $v$ to the root of $T$.  (This implies that $w=v$ when $v$ is a vertex of $F$.)  Since $w\in V(F)$, $w\in P_i$ for some $i\in\{1,\ldots,k\}$ and we define $\alpha(v):=i$.  Now add edges to the inner faces of $N'$ to obtain a supergraph $N''$ in which each inner face has three $3$ edges on its boundary. By Sperner's Lemma \note{DW}{Does this work for multigraph near-triangulations?} there exists an inner face $\tau'':=v_1,v_2,v_3$ of $N''$ such that $\alpha(v_i)=i$ for each $i\in\{1,2,3\}$.  Let $\tau:=v_1,\ldots,v_p$ be the face of $N'$ that contains $\tau''$.  (Note that $p\in\{3,4\}$.)

@@

%\referee{2}{Page 16.  At the end of Page 16, I do not see why 'every pair of edges cross at most once.'  Without further assumptions, it seems as if two paths in the path system could intersect several times.  Therefore, by drawing 'each edge vw of G alongside $P_{vw}$ in $G_0$', there may be two edges which cross several times.  It seems this can be fixed by choosing the path system carefully by 'uncrossing' paths which intersect at more than two internal vertices.}
%
%\note{DW}{I have revised the proof to address this error. I couldn't think of a way to do this without introducing balls and channels (and without vigorous hand-waving). Is it readable?}

@@

\citet{PS21} and \citet{DFMS21} use \cref{ps} and \cref{PlanarProduct}(b), respectively, to show that $\chi_p(G)$ is polynomial in $p$ when $G$ is planar or of bounded Euler genus.  
Upper bounds on $\chi_p$ for graphs of given treewidth~\citep{PS21} and for graph products~\citep{DFMS21} imply the next lemma.

%In the journal version, \citet{DFMS21} proved that $\chi_p(H_1\boxtimes H_2) \leq \chi_p(H_1) \cdot \chi(H_2^p)$ where $H_2^p$ is the $p$-th power of $H_2$. Apply this with $H_2:= P \boxtimes K_\ell$. Then $H_2^p= P^p\boxtimes K_\ell$ and $\chi(H_2^p)=(p+1)\ell$. So $\chi_p(H\boxtimes P \boxtimes K_\ell) \leq \chi_p(H_1) \cdot (p+1)\ell$. }

\begin{lem}[\citep{DFMS21,PS21}]
\label{p-centered}
For every graph $H$ of treewidth at most $t$ and for every path $P$, if $G\subseteq H\boxtimes P \boxtimes K_\ell$ then
\[\chi_p(G)\le \ell (p+1)\, \chi_p(H) \leq \ell (p+1) \tbinom{p+t}{t}.\]
%If $G\subseteq H\boxtimes P \boxtimes K_\ell$ then $\chi_p(G)\le \ell (p+1)\, \chi_p(H)$.
\end{lem}

%\note{DW}{For consistency with \cref{qn}, I suggest we combine \cref{p-centered,p-centered-treewidth}, and just write, ``}

%\begin{proof}
%	By \cref{PartitionProduct}, $G$ has an $H$-partition $(\mathcal{L}=\langle L_0,L_1,\ldots\rangle, \PP=(B_x:x\in V(H))$ with layered width at most $\ell$. Use a product colouring $\phi:V(G)\to \{1,\ldots,\ell\}\times\{0,\ldots,p\}\times\{1,\ldots,\chi_p(H)\}$.  For each integer $i\ge 0$ and each $x\in V(H)$, assign the colour $\phi(v):=(\alpha(v),\beta(v),\gamma(v))$ to each vertex $v\in L_i\cap B_x$ such that:
%   \begin{compactenum}
%     \item $\alpha(v)$ is unique among $\{\alpha(w): w\in L_i\cap B_x\}$, which is possible
%     since $|L_i\cap B_x|\le \ell$,
%     \item $\beta(v)= i\bmod (p+1)$, and
%     \item $\gamma(v)=\gamma(x)$ where $\gamma:V(H)\to\{1,\ldots,\chi_p(H)\}$ is a $p$-centered colouring of $H$.
%   \end{compactenum}
% To show this is a $p$-centered colouring, consider some connected subgraph $X\subseteq G$.
%
% First suppose that there exists $v,w\in V(X)$ with $v\in L_i$ and $w\in L_j$ with $j-i\ge p$. Since $X$ is connected, $X$ contains a path from $v$ to $w$.  By the definition of layering, this path contains at least one vertex from $L_{i'}$ for each $i'\in\{i,i+1,\ldots,j\}$. Therefore, $|\{\beta(v'):v'\in V(X)\}|\ge j-i+1 > p$, so $X$ receives more than $p$ distinct colours.
%
% Otherwise, $V(X)\subseteq L_{i}\cup\cdots\cup L_{i+s}$ for some $s<p$.  Let $H':=H[\{x\in V(H):B_x\cap V(X)\neq\emptyset]$.  If $|\{\gamma(x):x\in V(H')\}| > p$ then $|\{\gamma(v):v\in V(X)\}|> p$ so $|\{\phi(v):v\in V(X)\}|> p$ and we are done.  Otherwise, since $\gamma$ is a $p$-centered colouring of $H$, there must exist some $x\in V(H')$ such that $\gamma(x)\neq\gamma(y)$ for every $y\in V(H')\setminus\{x\}$.
% For any $v,w\in B_x$ with $v\neq w$, either $v,w\in L_{i'}$ for some $i'\in\{i,i+1,\ldots,i+s\}$ in which case $\alpha(v)\neq\alpha(w)$; or $v\in L_{i'}$ and $w\in L_{i''}$ with $|i'-i''|< p$, in which case $\beta(v)\neq\beta(w)$. Therefore every vertex $v\in B_x$ receives a colour $\phi(v)$ distinct from every colour in $\{\phi(z):z\in X\setminus\{x\}\}$. Therefore, every vertex in $B_x$ receives a colour distinct from every other vertex in $X$.
% \end{proof}
%
%\referee{2}{Page 18.  Replace '$|i'-i''| < p$' by  '$0< |i'-i''| < p$'.}

\cref{p-centered,kPlanarProduct,1-planar,gkPlanarProduct} immediately imply the following results, for every $p\ge 2$:

@@

Theorems~\ref{PlanarProduct}(b),  \ref{ShortcutProduct} and  \ref{GenusProduct}(b) and \cref{MapShortcut,qn,p-centered,non-repetitive} imply the following result.

@@

Theorems~\ref{PlanarProduct}(b), \ref{GenusProduct}(b) and \cref{ShortcutProduct} and  \cref{DrawG,qn,p-centered,non-repetitive,PowerShortcut} imply the following result, which with $g=0$ implies  \cref{kPowerBasic} in the introduction.

@@

\cref{qn,p-centered,non-repetitive,MinorFreeDegree,PowerProduct} imply the following analogous result for powers of graphs in any minor-closed class with bounded maximum degree.

@@

\cref{nearest-neighbour,kPlanarProduct,qn,p-centered} imply:

@@

\begin{lem}
	\label{NewLemma}
Let $G_0$ be a 2-connected plane multigraph. Let $T$ be a spanning tree of $G_0$ rooted at some vertex $r$ on the boundary of the outerface of $G_0$. Let $C = [P_1,\dots,P_k]$ be a cycle in $G_0$ with $r$ in the exterior of $C$, where $k\in\{1,2,3\}$ and $P_1, \dots, P_k$ are bipods in $G_0$. Let $G:=G_0\blah{C}$. Then $G$ has a partition $\PP$ into tripods with respect to $T$ such that $P_1,\dots,P_k\in\PP$, and $G / \PP$ has a tree-decomposition $\mathcal{T}$ of width at most 3 such that for every facial cycle $F$ of $G$ there is a bag $B$ of $\mathcal{T}$ such that every part of $\PP$ that intersects $F$ is in $B$.
\end{lem}

@@

First suppose that $C$ is a 2-cycle. Thus $k\leq 2$. Let $e$ be one of the edges in $C$. Let $F$ be the internal facial cycle of $G$ that includes $e$. Since each facial cycle of $G_0$ has length at least 3, $F$ consists of a non-empty path $P$ plus the edge $e$. Let $C'$ be obtained by replacing $e$ in $C$ by $P$. Let $G':=G\langle{C'}\rangle$. Thus $G'$ has one fewer face than $G$. Note that $C'=[P_1,\dots,P_k,P]$ (\note{DW}{or it can be written that way?}).  By induction, $G'$ has a partition $\PP$ into tripods with respect to $T$ such that $P_1,\dots,P_k,P\in\PP$, and $G / \PP$ has a tree-decomposition $\mathcal{T}$ of width at most 3 such that for every facial cycle $F'$ of $G$ there is a bag $B$ of $\mathcal{T}$ such that every part of $\PP$ that intersects $F'$ is in $B$. The bag of $\mathcal{T}$ that contains $P_1,\dots,P_k,P$ satisfies this property for $F$. Thus $\PP$ and $\TT$ satisfy the desired properties for $G$. Now assume that $|V(C)|\geq 3$.

@@

Consider $j\in\{1,\dots,m\}$. Let $G_j$ be the subgraph of $G$ consisting of $Z_j$ and all the edges and vertices of $G$ in the interior of $Z_j$. The outerface of $G_j$ is $Z_j$. Observe that for some $i\in\{1,2,3\}$, we have
$Z_j=[R_{i,j}^+,R_{i+1,j}^-,S_{i,j}]$, where 
$R_{i,j}^+$ is some subpath of $R_i^+$,
$R_{i+1,j}^-$ is some subpath of $R_{i+1}^-$, and
$L_{i,j}$ is the path $Z_j-V(C)$ (which is contained in $Y$). \note{DW}{Add a figure of a good case and a weird case.}
Since $R_i^+$, $R_{i+1}^-$ and $L_i$ are bipods, 
$R_{i,j}^+$, $R_{i+1,j}^-$ and $L_{i,j}$ are bipods. 
By construction, $v_{i-1}$ is not in $G_j$ (where $v_0=v_3$). 
Thus $V(G_j)\setminus V(Z_j) \subsetneq V(G)\setminus V(C)$, and we may apply induction to $G_j$.  Thus $G_j$ has a partition $\PP_j$ into tripods with respect to $T$ such that $R_{i,j}^+, R_{i+1,j}^-, L_{i,j}\in\PP_j$, and 	$G_j / \PP_j$ has a tree-decomposition  $\mathcal{T}_j=\{B_x:x\in V(D_j)\}$ of width at most 3 such that for every facial cycle $F'$ of $G_j$ there is a bag $B$ of $\mathcal{T}_j$ such that every part of $\PP_j$ that intersects $F'$ is in $B$.
